|

### 2024-09-05 manuscript/paperCMSACversion.Rmd
title change to match our submission

---
title: |
  |  \large Enforcers are not okay, probably.  
author: | 
abstract: |
 | A recent study looked at the relationship between long-term mortality and being an enforcer in the National Hockey League between 1967 and 2022.  They performed a matched cohort study that found no significant difference in the mortality rates of enforcers versus non-enforcers (where “enforcer” is defined based on number of career fights or average career penalty minutes).  They then perform an additional sub-analysis conditioning on players that were currently deceased and looking at just this subset of players, they find a significant difference in the mean age at death between the enforcers and non-enforcers groups (45.2 vs 55.2; p-value= 0.02) thus reaching the conclusion that enforcers died an average of 10 years earlier than non-enforcers. However, some of the statistical methods used to reach their conclusions were unconvincing, specifically the approach used to reach this last claim.  Therefore, we undertook a re-analysis of their study using a survival analysis approach, which we believe is the natural method to use to study this question.  In addition, we augmented their data by scraping hockey-reference to find more instances of players who were deceased.  Then, using a Cox proportional hazards model, we found no significant difference in survival rate between enforcers and non-enforcers when an enforcer was defined based on the number of career fights.  However, when an enforcer was defined by average career penalty minutes, a significant difference was observed, which was notably not found in the original article.  Specifically, we find an estimated hazard ratio of enforcers to non-enforcers of 2.29 (p-value: 0.00169).  Therefore, we agree with the sentiment of the original paper that there is some difference between survival of these two groups, but we believe our approach to estimating this effect is more appropriate and that the magnitude of the effect in the original article is overstated.
 \vspace{2mm} 
  |
 Keywords: survival analysis, hockey
bibliography: references.bib
fontsize: 10pt
link-citations: true
linkcolor: cyan
urlcolor: cyan
output:
  pdf_document:
    df_print: kable
    number_sections: true
    keep_tex: true
header-includes:
 \usepackage{setspace}
 \setstretch{1}
 \usepackage{float}
 \floatplacement{figure}{t}
---

```{r setup, echo=FALSE, message=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	fig.align = "center",
	message = FALSE,
	warning = FALSE
)

>>>>

---
title: |
  |  \large Assessing the Impact: A Survival (Re)Analysis of Hockey Enforcers vs. Non-Enforcers
author: | 
abstract: |
 | A recent study looked at the relationship between long-term mortality and being an enforcer in the National Hockey League between 1967 and 2022.  They performed a matched cohort study that found no significant difference in the mortality rates of enforcers versus non-enforcers (where “enforcer” is defined based on number of career fights or average career penalty minutes).  They then perform an additional sub-analysis conditioning on players that were currently deceased and looking at just this subset of players, they find a significant difference in the mean age at death between the enforcers and non-enforcers groups (45.2 vs 55.2; p-value= 0.02) thus reaching the conclusion that enforcers died an average of 10 years earlier than non-enforcers. However, some of the statistical methods used to reach their conclusions were unconvincing, specifically the approach used to reach this last claim.  Therefore, we undertook a re-analysis of their study using a survival analysis approach, which we believe is the natural method to use to study this question.  In addition, we augmented their data by scraping hockey-reference to find more instances of players who were deceased.  Then, using a Cox proportional hazards model, we found no significant difference in survival rate between enforcers and non-enforcers when an enforcer was defined based on the number of career fights.  However, when an enforcer was defined by average career penalty minutes, a significant difference was observed, which was notably not found in the original article.  Specifically, we find an estimated hazard ratio of enforcers to non-enforcers of 2.29 (p-value: 0.00169).  Therefore, we agree with the sentiment of the original paper that there is some difference between survival of these two groups, but we believe our approach to estimating this effect is more appropriate and that the magnitude of the effect in the original article is overstated.
 \vspace{2mm} 
  |
 Keywords: survival analysis, hockey
bibliography: references.bib
fontsize: 10pt
link-citations: true
linkcolor: cyan
urlcolor: cyan
output:
  pdf_document:
    df_print: kable
    number_sections: true
    keep_tex: true
header-includes:
 \usepackage{setspace}
 \setstretch{1}
 \usepackage{float}
 \floatplacement{figure}{t}
---

<!-- old title: Enforcers are not okay, probably. -->
```{r setup, echo=FALSE, message=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	fig.align = "center",
	message = FALSE,
	warning = FALSE
)
